Developing a Bubble Chamber Particle Discriminator Using Semi-Supervised Learning

### 2018-11-13 documentation/paper.tex
Lots of general edits to the paper

\usepackage[margin=1in]{geometry}
\usepackage{textcomp}
\usepackage{parskip}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage[backend=biber, style=alphabetic, sorting=ynt]{biblatex}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{caption}
\usepackage{lineno}
\linenumbers

@@

@@

In the PICO-60 experiment, and in most other experiments striving to directly detect dark matter, one of the most important and challenging hurdles to overcome is that of background events. Unwanted particles from a variety of sources can produce event signatures that appear very similar to those which are expected to be created by dark matter candidates.

@@

    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only run types as training data. While run types are known to be mostly correlated with particle types, the relatively large percentage of impurities has the potential to hinder conventional learning techniques. Specifically, the run types available for training are:

@@

The architecture described may have wider applications to several similar cases, such as those that require a discriminator which is difficult to construct using conventional techniques.

@@

The PICO-60 bubble chamber experiment, created for the detection of weakly interacting massive particles (WIMPs) \cite{pico}, was completed in 2017. It provides an excellent basis on which to develop and test the efficacy of new machine learning discriminators, relative to the techniques used during initial analysis of the experiment.

@@

Alpha decays along the U/Th chains present the dominant remaining background which must be distinguished from nuclear recoils acoustically.

@@

WIMP candidates in the cross-section and mass ranges to which the PICO-60 detector is sensitive are predicted to produce nuclear recoils with the same properties as those induced by neutrons. The key difference is that neutrons frequently (but not always) scatter several times and produce multiple bubbles, where the extremely small predicted cross-sections of WIMP candidates mean that they should almost invariably create just one bubble. This means single-bubble neutron events can be used to optimize a discriminator to detect WIMP events.

The approximate expected ratio between single-bubble and multiple-bubble events generated by neutrons is known. A measured quantity of single-bubble events generated by nuclear recoils, in excess of this prediction, would be indicative of WIMP interactions. Consequently, it is imperative to isolate only those events associated with nuclear recoils.

@@

Before AP can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$). It must also be preprocessed with the position correction function $PosCor(\beta_{8}, \chi)$, which adjusts the overall volume of an audio recording according to its position in the vessel. This is done because the acoustic characteristics differ significantly depending on the position of the event in the vessel.

Finally, a set of data cuts must be applied to remove classes of events which are not accurately classified using AP alone. These classes include wall event and multiple-bubble events.

@@

The lowest-level data derived from the piezos is the raw waveform $\omega$. This consists of a series of samples, collected at the rate of one per microsecond. They are each represented as a 16-bit integer. Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because there is only background noise prior to this window, and a clipped signal produced by hydraulics repressurizing the vessel after. These sections contain no information and can be undesirably fit on by neural networks. (For context, there are 250,000 samples in the entire waveform, and approximately half of them are recorded prior to the camera trigger.)

@@

\subsubsection{Camera-Derived}

@@

For each event that is captured, four cameras within the PICO-60 apparatus each capture a sequence of 41 images before and after formation of a bubble is detected using the cameras and an image-based trigger. These raw images contain a large amount of extraneous information; they encompass the entire vessel. To reduce the input information, the image window sequence $\iota$ includes 50\texttimes50 cropped windows around the position of the bubble. The 41 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble. Five frames from before the recording trigger are included in $\iota$, because the bubble does not cause a trigger until it is already of a significant size.

@@

The 3-dimensional position $\chi$ of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera. This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble. It is also used for {\it FidWallCut}$(D)$.

@@

PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles and WIMP candidates never multiply scatter. Thus, no discriminator is required to handle these events, so they are removed. The bubble multiplicity cut {\it MultiCut}$(D)$ consists of the following restrictions:

@@

Events that occur near the walls of the vessel have acoustic properties which are notably different from events nearer the center of the vessel. It can be desirable for a discriminator to handle these events correctly; however, AP alone does not, and neither does the neural network used in the previous PICO-60 paper. Thus, removing wall events allowed for a more meaningful direct comparison between a new neural network and existing techniques.

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{wall_event_positions}
    \caption{\label{} A visualization of the fiducial cuts and how they match up with the pressure and acoustic cuts.}
\end{figure}

@@

@@

While machine learning has existed for many years, recent developments in artificial neural networks have opened up a wide variety of applications and fields in which it can now be used.

@@

This characteristic has already shown wide-ranging implications in many fields. For one, it has the potential to drastically increase the speed of iteration for the people working on a project, since retraining a machine learning system when a hardware or software variable changes is much faster than calibrating a human-designed model.

@@

Two original semi-supervised algorithms were developed and implemented for this study: gravitational differentiation and iterative cluster nucleation. Both are discussed in detail in Section \ref{semi_supervised}.

@@

$S=std(N \cup A)$

$CWSD=(std(N \div S) + std(A \div S)) \div 2$

@@

The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually. This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are for each class. Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any $x$ and $y$), with minimal variance off those specific values, will produce a low class-wise standard deviation.

@@

The following tables show the results of several {\it DeepConv}$(\omega)$ discriminators produced throughout testing, with corresponding network hyperparameters.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of {\it DeepConv}$(\omega)$ configurations.}
        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
            \hline
            L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Max Accuracy & Precision & Recall & CWSD \\
            \hline
            0.003 & 0 & 24 & 3 & 17 & 95\% & 97\% & 97\% & 0.52 \\
            % ../experimental_data/validation_results/waveform_grid_search/l2_lambda0.003/dense_dropout0/first_layer_filters24/kernel_size3/convolutional_layers_per_group3/time1539990193_epoch57.json
            \hline
            0.003 & 0.25 & 48 & 3 & 17 & 94\% & 100\% & 91\% & 0.44 \\
            \hline
            0.001 & 0.25 & 24 & 5 & 17 & 95\% & 99\% & 94\% & 0.46 \\
            \hline
            0.001 & 0.25 & 24 & 5 & 32 & 91\% & 94\% & 94\% & 0.66 \\
            \hline
            0.0003 & 0 & 48 & 5 & 32 & 88\% & 94\% & 89\% & 0.61 \\
            \hline
            0.0003 & 0 & 24 & 3 & 17 & 93\% & 96\% & 95\% & 0.59 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

The following tables compare various multi-layer perceptron discriminators, with and without wall cuts.

@@

It is very clear in Table \ref{fft_wall_cuts} that higher resolution data does not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.

@@

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of {\it ImageConv}$(\iota)$ configurations.}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            L2 $\lambda$ & Dropout & Convolutional Layers & Dense Layers & Max Accuracy & Precision & Recall & CWSD \\
            \hline
            0 & 0 & 6 & 3 & 69\% & 69\% & 100\% & 0.99 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0_dropout0_convolutional_layers_per_group2time1532891695_epoch15.json 
            \hline
            0.01 & 0.25 & 6 & 3 & 68\% & 68\% & 100\% & 0.98 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group2time1532895825_epoch2.json
            \hline
            0.003 & 0.25 & 9 & 3 & 68\% & 68\% & 100\% & 0.98 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.003_dropout0.25_convolutional_layers_per_group3time1532895118_epoch3.json
            \hline
            0.01 & 0.25 & 12 & 3 & 68\% & 68\% & 100\% & 0.97 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group4time1532896126_epoch0.json
            \hline
            0.01 & 0.5 & 12 & 3 & 68\% & 68\% & 100\% & 0.99 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.5_convolutional_layers_per_group4time1532898944_epoch1.json
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

The high standard deviations across the board are further evidence for the hypothesis that there is insufficient information in the image data. The 100\% recall statistics are not due to high classification accuracy, but rather due to the network making predictions less than 0.5 for all examples of both classes.

@@

The following tables compare the accuracy and mean class-wise standard deviation values of certain successful supervised learning configurations with multiple input formats.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Summary of supervised learning configurations.}
        \begin{tabular}{|l|l|l|l|l|}
            \hline
            Configuration & Max Accuracy & Precision & Recall & CWSD \\
            \hline
            {\it DeepConv}$(\omega)$ & 95\% & 97\% & 97\% & 0.52 \\
            \hline
            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}))$ & 96\% & 98\% & 97\% & 0.42 \\
            \hline
            {\it FourierMLP}$(\beta_{8})$ & 98\% & 100\% & 98\% & 0.29 \\
            \hline
            {\it FourierMLP}$(\beta_{50,001})$ & 95\% & 95\% & 98\% & 0.46 \\
            \hline
            {\it ImageConv}$(\iota)$ & 68\% & 68\% & 100\% & 0.97 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

The key conclusion to be drawn from the above tables is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format. Higher resolutions seem to make performance worse, likely due to overfitting.

@@

Gradients are calculated using the gravitational differentiation function $GravDiff(p, \psi, g)$, which is parameterized by the network's existing prediction $p$ on the training example in question, the degree $\psi$ of the piecewise exponential function used to distort the response of the gradient, and the gravitational multiplier $g$. The function is defined as follows:

{\it GravDiff}$(p, \psi, g) = g \cdot sgn(p) \cdot abs(tanh(2(p - 0.5))) ^ \psi$

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{grav_diff}
    \caption{\label{grav_diff} A visualization of {\it GravDiff} $(p, \psi, g)$ with respect to $p$ and $\psi$.}
\end{figure}

@@

\begin{enumerate}
    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate {\it GravDiff}$(p, \psi, g)$ and record the resulting gradients for the next iteration.
    \item Increment $g$ by $\delta_{g}$. At the beginning, when $g = 0$, training will be entirely based on the labeled set $\varsigma$, and progressively, over the course of a run, the gravitational effect increases with $g$.
\end{enumerate}

@@

The following table represents the performance of several high-accuracy gravitational differentiation configurations tested during the grid search. The accuracy statistic is the mean over three tests of the same configuration.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of gravitational differentiation configurations.}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Mean AP Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 0.0005 & 0.001 & 5 & 73\% & 72\% & 90\% & 0.78 \\
            \hline
            32 & 0.0005 & 0.003 & 9 & 83\% & 80\% & 99\% & 0.63 \\
            \hline
            64 & 0.0005 & 0.003 & 9 & 77\% & 74\% & 97\% & 0.71 \\
            \hline
            64 & 0.003 & 0.03 & 7 & 97\% & 96\% & 100\% & 0.18 \\
            \hline
            128 & 0.008 & 0.001 & 9 & 74\% & 71\% & 95\% & 0.70 \\
            \hline
            128 & 0.0005 & 0.03 & 11 & 99\% & 99\% & 99\% & 0.17 \\
            \hline
            256 & 0.005 & 0.01 & 11 & 97\% & 96\% & 99\% & 0.25 \\
            \hline
            256 & 0.001 & 0.03 & 3 & 99\% & 99\% & 100\% & 0.10 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

The following table represents the performance of a variety of high-accuracy iterative cluster nucleation configurations tested during the grid search. As with gravitational differentiation, the accuracy statistic is the mean over three tests of the same configuration.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of iterative cluster nucleation configurations.}
        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
            \hline
            Initial $\varsigma$ Size & L2 $\lambda$ & Dropout & Initial $j$ & $k$ & Mean AP Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 0 & 0.25 & 0.01 & 1.025 & 80\% & 79\% & 99\% & 0.60 \\
            \hline
            32 & 0 & 0 & 0.01 & 1.025 & 95\% & 94\% & 99\% & 0.38 \\
            \hline
            64 & 0.001 & 0 & 0.01 & 1.05 & 97\% & 96\% & 99\% & 0.28 \\
            \hline
            64 & 0 & 0.25 & 0.02 & 1.025 & 97\% & 95\% & 100\% & 0.27 \\
            \hline
            128 & 0 & 0.5 & 0.01 & 1.05 & 92\% & 90\% & 100\% & 0.35 \\
            \hline
            128 & 0 & 0.5 & 0.02 & 1.05 & 98\% & 98\% & 100\% & 0.21 \\
            \hline
            256 & 0 & 0.5 & 0.01 & 1.05 & 99\% & 99\% & 100\% & 0.15 \\
            \hline
            256 & 0.003 & 0.5 & 0.01 & 1.05 & 98\% & 98\% & 98\% & 0.26 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

The following table directly compares performance of the most successful semi-supervised learning configurations, with AP provided as a baseline for comparison. Remember that ``Mean AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Summary of the most successful semi-supervised learning configurations.}
        \begin{tabular}{|l|l|l|l|l|l|}
            \hline
            Initial $\varsigma$ Size & Technique & Mean AP Accuracy & Mean Precision & Mean Recall & Mean CWSD \\
            \hline
            32 & GD & 83\% & 80\% & 99\% & 0.63 \\
            \hline
            32 & ICN & 95\% & 94\% & 99\% & 0.38 \\
            \hline
            64 & GD & 97\% & 96\% & 100\% & 0.18 \\
            \hline
            64 & ICN & 97\% & 96\% & 99\% & 0.28 \\
            \hline
            128 & GD & 99\% & 99\% & 99\% & 0.17 \\
            \hline
            128 & ICN & 98\% & 98\% & 100\% & 0.21 \\
            \hline
            256 & GD & 99\% & 99\% & 100\% & 0.10 \\
            \hline
            256 & ICN & 99\% & 99\% & 100\% & 0.15 \\
            \hline
            278 & PICO-60 ML & 80\% & 76\% & 100\% & 0.64 \\
            \hline
            $\infty$ & AP & 100\% & 100\% & 100\% & 0.41 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

The following table compares the overall average performance of each of the semi-supervised learning configurations for different labeled set sizes.

>>>>

\usepackage[margin=1in]{geometry}
\usepackage{textcomp}
\usepackage{parskip}
\usepackage{amsmath}
\usepackage{authblk}
\usepackage[backend=biber, style=alphabetic, sorting=ynt]{biblatex}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{lineno}
\linenumbers

@@

\setcounter{section}{-1}

\section{Structure}

This paper is organized as follows:
\begin{itemize}
    \item In Section 1, we outline generally the PICO-60 experiment and the context in which machine learning is applicable.
    \item In Section 2, we discuss the specific properties of the experiment and its data, as well as the existing discriminator.
    \item In Section 3, we discuss various machine learning techniques that are applied to develop a new discriminator.
    \item In Section 4, we document experimentation with supervised learning, applied to various data formats.
    \item In Section 5, we document experimentation with semi-supervised learning with the goal of improving data efficiency.
    \item In Section 6, we conclude this study and discuss its implications.
    \item In Section 7, we overview the technologies and software used for this study.
\end{itemize}

@@

In the PICO-60 \cite{pico} experiment, and in most other experiments striving to directly detect WIMP dark matter \cite{wimp}, one of the most important and challenging hurdles to overcome is that of background events. Unwanted particles from a variety of sources can produce event signatures that appear very similar to those which are expected to be created by dark matter candidates.

@@

    For clarity, and with specific reference to the PICO-60 experiment, the intent is to accurately separate particle types using only calibration/background data collection run types (henceforth "run types") as training data. While run types are known to be mostly correlated with particle types, the relatively large percentage of impurities has the potential to hinder conventional learning techniques. Specifically, the run types available for training are:

@@

The architectures described above may have wider applications to several similar cases, such as those that require a discriminator which is difficult to construct using conventional techniques.

@@

The PICO-60 \cite{pico} bubble chamber experiment, created for the detection of weakly interacting massive particles (WIMPs), was completed in 2017. It provides an excellent basis on which to develop and test the efficacy of new machine learning discriminators, relative to the techniques used during initial analysis of the experiment.

@@

Alpha decays along the U/Th chains present the dominant remaining background which must be distinguished from nuclear recoils acoustically \cite{isotopes}.

@@

The nuclear recoils produced by WIMP candidates (in the cross-section and mass ranges to which the PICO-60 detector is sensitive) are predicted to be indistinguishable from those induced by neutrons. The key difference is that neutrons frequently (but not always) scatter several times and produce multiple bubbles, where the extremely small predicted cross-sections of WIMP candidates mean that they should almost invariably create just one bubble. This means single-bubble neutron events can be used to optimize a discriminator to detect WIMP events.

The approximate expected ratio between single-bubble and multiple-bubble events generated by neutrons is known based on Monte Carlo simulation and extrapolations from experience with previous detectors. A measured quantity of single-bubble events generated by nuclear recoils, in excess of this prediction, would be indicative of WIMP interactions. Consequently, it is imperative to isolate only those events associated with nuclear recoils.

@@

Before AP can be calculated, the audio data must first be converted into the frequency domain (specifically, with the banded Fourier transform $\beta_{8}$). It must also be preprocessed with the position correction function (discussed in Section \ref{camera_derived}), which adjusts the overall volume of an audio recording according to its position in the vessel. This is done because the acoustic characteristics differ significantly depending on the position of the event in the vessel \cite{pico}.

Finally, a set of data cuts must be applied to remove classes of events which are not accurately classified using AP alone. These classes include events originating close to the walls of the detector, and multiple-bubble events.

@@

The lowest-level data derived from the piezos is the raw waveform $\omega$. This consists of a series of 250,000 samples, collected at the rate of one per microsecond. Each sample is represented as a 16-bit integer. Only the section of the audio between sample 90,000 (inclusive) and sample 190,000 (exclusive) was used during most experiments, because there is only background noise prior to this window, and a clipped signal produced by hydraulics repressurizing the vessel after. These sections contain no information and can be undesirably fit on by neural networks. (For context, approximately half of the samples are recorded prior to the camera trigger.)

@@

\subsubsection{Camera-Derived} \label{camera_derived}

@@

Bubble events are detected using the four cameras within the PICO-60 apparatus, with an image-based trigger. When this happens, each camera records a sequence of 41 images before and after formation. These raw images contain a large amount of extraneous information; they encompass the entire vessel. To reduce the input information, the image window sequence $\iota$ includes 50\texttimes50 cropped windows around the position of the bubble. The 41 frames, many of which contain either no bubble or a bubble in later stages of formation, are reduced to 10 immediately around the formation of the bubble. Five frames from before the recording trigger are included in $\iota$, because the bubble does not cause a trigger until it is already of a significant size.

@@

The 3-dimensional position $\chi$ of the bubble within the vessel is calculated using triangulation, based on the known positions and angles of the cameras and the position of the bubble within the field of view of each camera. This is used in the banded frequency position correction function $PosCor(\beta _{8}, \chi)$, which corrects $\beta _{8}$ for variations in amplitude which depend on the position of the bubble.

@@

PICO-60 events which include multiple bubbles are \textit{always} neutron events; alpha particles are stopped electromagnetically, and WIMP candidates have a negligible probability of multiply scattering because of their extremely low predicted cross section. Thus, no discriminator is required to handle these events, so they are removed. The bubble multiplicity cut {\it MultiCut}$(D)$ consists of the following restrictions:

@@

Events that occur near the walls of the vessel have acoustic properties which are very different from events nearer the center of the vessel. It can be desirable for a discriminator to handle these events correctly; however, AP alone does not, and neither does the neural network used in the previous PICO-60 paper. Thus, removing wall events allowed for a more meaningful direct comparison between a new neural network and existing techniques.

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{wall_event_positions}
    \caption{\label{wall_cuts_visualization} A visualization of the fiducial cuts and how they match up with the pressure and acoustic cuts.}
\end{figure}

@@

In Figure \ref{wall_cuts_visualization}, the relative and combined effectiveness of each of these cuts can be seen. It is clear that most of the time, both of these cuts agree, but there are certain cases in which only one or the other marks an event.

@@

While neural networks have existed for many years, recent developments in computing power, most importantly GPU acceleration \cite{gpu}, have opened up a wide variety of applications and fields in which they can now be used.

@@

This characteristic has already shown wide-ranging implications in many fields, such as autonomous driving \cite{end-to-end} and detection of heart arrhythmia \cite{arrhythmia}. For one, it has the potential to drastically increase the speed of iteration for the people working on a project, since retraining a machine learning system when a hardware or software variable changes is much faster than calibrating a human-designed model.

@@

Two original semi-supervised algorithms were developed and implemented for this study: gravitational differentiation (GD) and iterative cluster nucleation (ICN). Both are discussed in detail in Section \ref{semi_supervised}.

@@

\begin{equation}
    S=std(N \cup A)
\end{equation}
\begin{equation}
    \mbox{\it CWSD}=(std(N \div S) + std(A \div S)) \div 2
\end{equation}

@@

The second step is to calculate the mean of the standard deviations of the normalized sets of neutrons and alpha particles individually. This is an indication of how tightly clustered or widely dispersed the discriminator’s predictions are for each class. Very consistent predictions of $x$ for neutrons and $y$ for alphas (for any given $x$ and $y$), with minimal variance off those specific values, will produce a low class-wise standard deviation.

@@

Table \ref{deepconv_table} shows the results of several {\it DeepConv}$(\omega)$ discriminators produced throughout testing, with corresponding network hyperparameters.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of {\it DeepConv}$(\omega)$ configurations.} \label{deepconv_table} 
        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
            \hline
            L2 $\lambda$ & Dropout & Filters & Kernel & Conv Layers & Max Accuracy & Precision & Recall & CWSD \\
            \hline
            0.003 & 0 & 24 & 3 & 17 & 95\% & 97\% & 97\% & 0.52 \\
            % ../experimental_data/validation_results/waveform_grid_search/l2_lambda0.003/dense_dropout0/first_layer_filters24/kernel_size3/convolutional_layers_per_group3/time1539990193_epoch57.json
            \hline
            0.003 & 0.25 & 48 & 3 & 17 & 94\% & 100\% & 91\% & 0.44 \\
            \hline
            0.001 & 0.25 & 24 & 5 & 17 & 95\% & 99\% & 94\% & 0.46 \\
            \hline
            0.001 & 0.25 & 24 & 5 & 32 & 91\% & 94\% & 94\% & 0.66 \\
            \hline
            0.0003 & 0 & 48 & 5 & 32 & 88\% & 94\% & 89\% & 0.61 \\
            \hline
            0.0003 & 0 & 24 & 3 & 17 & 93\% & 96\% & 95\% & 0.59 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

Table \ref{fft_wall_cuts} compares various multi-layer perceptron discriminators, with and without wall cuts.

@@

It is very clear that higher resolution data does not improve accuracy. Regularization, both dropout and L2, appears to have relatively little effect.

@@

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of {\it ImageConv}$(\iota)$ configurations.} \label{imageconv_table}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            L2 $\lambda$ & Dropout & Convolutional Layers & Dense Layers & Max Accuracy & Precision & Recall & CWSD \\
            \hline
            0 & 0 & 6 & 3 & 69\% & 69\% & 100\% & 0.99 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0_dropout0_convolutional_layers_per_group2time1532891695_epoch15.json 
            \hline
            0.01 & 0.25 & 6 & 3 & 68\% & 68\% & 100\% & 0.98 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group2time1532895825_epoch2.json
            \hline
            0.003 & 0.25 & 9 & 3 & 68\% & 68\% & 100\% & 0.98 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.003_dropout0.25_convolutional_layers_per_group3time1532895118_epoch3.json
            \hline
            0.01 & 0.25 & 12 & 3 & 68\% & 68\% & 100\% & 0.97 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.25_convolutional_layers_per_group4time1532896126_epoch0.json
            \hline
            0.01 & 0.5 & 12 & 3 & 68\% & 68\% & 100\% & 0.99 \\
            % ../experimental_data/validation_results/2018_07_29_image_grid_search/image_grid_search_l2_lambda0.01_dropout0.5_convolutional_layers_per_group4time1532898944_epoch1.json
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

In Table \ref{imageconv_table}, high standard deviations across the board are further evidence for the hypothesis that there is insufficient information in the image data. The 100\% recall statistics are not due to high classification accuracy, but rather due to the network making predictions less than 0.5 for all examples of both classes.

@@

Table \ref{super_overview} compares the accuracy and mean class-wise standard deviation values of certain successful supervised learning configurations with multiple input formats.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Summary of supervised learning configurations.} \label{super_overview}
        \begin{tabular}{|l|l|l|l|l|}
            \hline
            Configuration & Max Accuracy & Precision & Recall & CWSD \\
            \hline
            {\it DeepConv}$(\omega)$ & 95\% & 97\% & 97\% & 0.52 \\
            \hline
            {\it FourierMLP}$(${\it PosCor}$(\beta_{8}))$ & 96\% & 98\% & 97\% & 0.42 \\
            \hline
            {\it FourierMLP}$(\beta_{8})$ & 98\% & 100\% & 98\% & 0.29 \\
            \hline
            {\it FourierMLP}$(\beta_{50,001})$ & 95\% & 95\% & 98\% & 0.46 \\
            \hline
            {\it ImageConv}$(\iota)$ & 68\% & 68\% & 100\% & 0.97 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

The key conclusion to be drawn from the above table is that the low-resolution Fourier transform $\beta_{8}$ is by far the most effective input format. Higher resolutions seem to make performance worse, likely due to overfitting.

@@

Gradients are calculated using the gravitational differentiation function $\mathrm{GravDiff}(p, \psi, g)$, which is parameterized by the network's existing prediction $p$ on the training example in question, the degree $\psi$ of the piecewise exponential function used to distort the response of the gradient, and the gravitational multiplier $g$. The function is defined as follows:

\begin{equation}
    \mathrm{GravDiff}(p, \psi, g) = g \cdot \mathrm{sgn}(p) \cdot \mathrm{abs}(\mathrm{tanh}(2(p - 0.5))) ^ \psi
\end{equation}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{grav_diff}
    \caption{\label{grav_diff} A visualization of $\mathrm{GravDiff}(p, \psi, g)$ with respect to $p$ and $\psi$.}
\end{figure}

@@

\begin{enumerate}
    \item Train $NN(x)$ for a single epoch on the combined set $\varsigma \cup \upsilon$, using the predefined ground truths for $\varsigma$ (with a mean squared error loss function) and the most recently calculated gravitational gradients for each of the examples in $\upsilon$.
    \item Calculate predictions $p$ with $NN(x)$ on the entirety of $\upsilon$. Calculate $\mathrm{GravDiff}(p, \psi, g)$ and record the resulting gradients for the next iteration.
    \item Increment $g$ by $\delta_{g}$. At the beginning, when $g = 0$, training will be entirely based on the labeled set $\varsigma$, and progressively, over the course of a run, the gravitational effect increases with $g$.
\end{enumerate}

@@

Table \ref{grav_overview} represents the performance of several high-accuracy gravitational differentiation configurations tested during the grid search. The accuracy statistic is the mean over three tests of the same configuration.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of gravitational differentiation configurations.} \label{grav_overview}
        \begin{tabular}{|l|l|l|l|l|l|l|l|}
            \hline
            $\varsigma$ Size & $\delta g$ & Learning Rate & $\psi$ & Mean AP Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 0.0005 & 0.001 & 5 & 73\% & 72\% & 90\% & 0.78 \\
            \hline
            32 & 0.0005 & 0.003 & 9 & 83\% & 80\% & 99\% & 0.63 \\
            \hline
            64 & 0.0005 & 0.003 & 9 & 77\% & 74\% & 97\% & 0.71 \\
            \hline
            64 & 0.003 & 0.03 & 7 & 97\% & 96\% & 100\% & 0.18 \\
            \hline
            128 & 0.008 & 0.001 & 9 & 74\% & 71\% & 95\% & 0.70 \\
            \hline
            128 & 0.0005 & 0.03 & 11 & 99\% & 99\% & 99\% & 0.17 \\
            \hline
            256 & 0.005 & 0.01 & 11 & 97\% & 96\% & 99\% & 0.25 \\
            \hline
            256 & 0.001 & 0.03 & 3 & 99\% & 99\% & 100\% & 0.10 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

Table \ref{icn_overview} represents the performance of a variety of high-accuracy iterative cluster nucleation configurations tested during the grid search. As with gravitational differentiation, the accuracy statistic is the mean over three tests of the same configuration.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Performance of iterative cluster nucleation configurations.} \label{icn_overview}
        \begin{tabular}{|l|l|l|l|l|l|l|l|l|}
            \hline
            Initial $\varsigma$ Size & L2 $\lambda$ & Dropout & Initial $j$ & $k$ & Mean AP Accuracy & Precision & Recall & CWSD \\
            \hline
            32 & 0 & 0.25 & 0.01 & 1.025 & 80\% & 79\% & 99\% & 0.60 \\
            \hline
            32 & 0 & 0 & 0.01 & 1.025 & 95\% & 94\% & 99\% & 0.38 \\
            \hline
            64 & 0.001 & 0 & 0.01 & 1.05 & 97\% & 96\% & 99\% & 0.28 \\
            \hline
            64 & 0 & 0.25 & 0.02 & 1.025 & 97\% & 95\% & 100\% & 0.27 \\
            \hline
            128 & 0 & 0.5 & 0.01 & 1.05 & 92\% & 90\% & 100\% & 0.35 \\
            \hline
            128 & 0 & 0.5 & 0.02 & 1.05 & 98\% & 98\% & 100\% & 0.21 \\
            \hline
            256 & 0 & 0.5 & 0.01 & 1.05 & 99\% & 99\% & 100\% & 0.15 \\
            \hline
            256 & 0.003 & 0.5 & 0.01 & 1.05 & 98\% & 98\% & 98\% & 0.26 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

Table \ref{semi_success_overview} directly compares performance of the most successful semi-supervised learning configurations, with AP provided as a baseline for comparison. Remember that ``Mean AP Accuracy'' refers to accuracy at replication of AP, rather than replication of the ground truths.

\begin{minipage}{\textwidth}
    \begin{center}
        \captionof{table}{Summary of the most successful semi-supervised learning configurations.} \label{semi_success_overview}
        \begin{tabular}{|l|l|l|l|l|l|}
            \hline
            Initial $\varsigma$ Size & Technique & Mean AP Accuracy & Mean Precision & Mean Recall & Mean CWSD \\
            \hline
            32 & GD & 83\% & 80\% & 99\% & 0.63 \\
            \hline
            32 & ICN & 95\% & 94\% & 99\% & 0.38 \\
            \hline
            64 & GD & 97\% & 96\% & 100\% & 0.18 \\
            \hline
            64 & ICN & 97\% & 96\% & 99\% & 0.28 \\
            \hline
            128 & GD & 99\% & 99\% & 99\% & 0.17 \\
            \hline
            128 & ICN & 98\% & 98\% & 100\% & 0.21 \\
            \hline
            256 & GD & 99\% & 99\% & 100\% & 0.10 \\
            \hline
            256 & ICN & 99\% & 99\% & 100\% & 0.15 \\
            \hline
            278 & PICO-60 ML & 80\% & 76\% & 100\% & 0.64 \\
            \hline
            $\infty$ & AP & 100\% & 100\% & 100\% & 0.41 \\
            \hline
        \end{tabular}
    \end{center}
\end{minipage}

@@

Table \ref{average_semi} compares the overall average performance of each of the semi-supervised learning configurations for different labeled set sizes.
